
```{r}
library(tidyverse)
library(DECIPHER)
library(bioseq)
library(Biostrings)
library(ape)
```

```{r}
read_fasta_df <- function (file = "") {
  fasta <- readLines(file)
  ind <- grep(">", fasta)
  s <- data.frame(ind = ind, from = ind + 1, to = c((ind - 
    1)[-1], length(fasta)))
  seqs <- rep(NA, length(ind))
  for (i in 1:length(ind)) {
    seqs[i] <- paste(fasta[s$from[i]:s$to[i]], collapse = "")
  }
  tib <- tibble(label = gsub(">", "", fasta[ind]), sequence = seqs)
  return(tib)
}


write_fasta <- function(sequences, names, filename) {
  con <- file(filename, "w")
  for (i in seq_along(sequences)) {
    writeLines(paste0(">", names[i]), con)
    if (!is.character(sequences[[i]])) {
      sequences[[i]] <- as.character(sequences[[i]])
    }
    writeLines(sequences[[i]], con)
  }
  close(con)
}


write_fasta_df <- function (data, filename) 
{
    fastaLines = c()
    for (rowNum in 1:nrow(data)) {
        fastaLines = c(fastaLines, as.character(paste(">", 
            data[rowNum, "label"], sep = "")))
        fastaLines = c(fastaLines, as.character(data[rowNum, 
            "sequence"]))
    }
    fileConn <- file(filename)
    writeLines(fastaLines, fileConn)
    close(fileConn)
}

dna_to_DNAbin <- function (dna){
  DNAbin <- as_DNAbin(dna)
  names(DNAbin) <- names(dna)
  return(DNAbin)
}
dna_to_DNAStringset <- function(x) 
{
    bioseq:::check_dna(x)
    DNAstr <- DNAStringSet(paste(x))
    names(DNAstr) <- names(x)
    return(DNAstr)
}

DNAStringSet_to_dna <- function(x){
    x_dna <- as_dna(paste(x))
    names(x_dna) <- names(x)
    res <- tibble(label = names(x), sequence = x_dna)
    return(res)
}

# Convert DNAstringset to DNAbin
DNAStringSet_to_DNAbin <- function(DNAStringSet){
  DNAbin <- as.DNAbin(DNAStringSet)
  return(DNAbin)
}

# https://bmcbioinformatics.biomedcentral.com/articles/10.1186/s12859-018-2315-y
palette <- c("A" = "#46ff2d", 
             "G" = "#ffae01", 
             "C" = "#f24641", 
             "T" = "#4294fa", 
             "K" = "#8b4816",
             "M" = "#83831f",
             "R" = "#ffff81",
             "S" = "#ff9d80",
             "Y" = "#e381f2",
             "W" = "#80fff2",
             "V" = "#fde4b8",
             "B" = "#f9c1bf",
             "H" = "#c0d9f9",
             "D" = "#c7ffba",
             "U" = "#8989fb",
             "N" = "black", 
             "-" = "white",
             "+" = "deeppink")


pal_df <- data.frame(names = names(palette), col = palette)


```

```{r}
#psba txt files forward and reverse
f_all <- sort(list.files(path = "/Users/bross/Desktop/AIMS/Analysis/sanger_analysis/sanger_raw", pattern = "F.txt", full.names = TRUE))

r_all <- sort(list.files(path = "/Users/bross/Desktop/AIMS/Analysis/sanger_analysis/sanger_raw", pattern = "R.txt", full.names = TRUE))

f_all
r_all

f_all_fasta <- f_all %>%
  map(read_fasta_df) %>% # read in all the files
  purrr::reduce(rbind) # reduce with rbind into one dataframe

r_all_fasta <- r_all %>%
  map(read_fasta_df) %>% # read in all the files
  purrr::reduce(rbind) # reduce with rbind into one dataframe
```

```{r}

# convert to format for DECIPHER and reverse complement the reverse sequence
f_all_fasta_sub <- f_all_fasta %>%
  mutate(sequence = case_when(str_detect(label, "49-1_Cob") ~ str_sub(sequence, start = 30, end = 890),
                              str_detect(label, "55-1_Cob") ~ str_sub(sequence, start = 30, end = 890),
                              str_detect(label, "49-2_Cob") ~ str_sub(sequence, start = 30, end = 890),
                              str_detect(label, "49-3_Cob") ~ str_sub(sequence, start = 30, end = 890),
                              str_detect(label, "49-1_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "49-2_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "49-3_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "55-1_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "55-4_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "55-5_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "49-1_cp23S") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "49-2_cp23S") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "49-3_cp23S") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "55-1_cp23S") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "49-1_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "49-2_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "49-3_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "55-1_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "55-4_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "55-5_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              TRUE ~ str_sub(sequence, start = 30, end = 600)))


r_all_fasta_sub <- r_all_fasta %>%
  mutate(sequence = case_when(str_detect(label, "49-1_Cob") ~ str_sub(sequence, start = 30, end = 890),
                              str_detect(label, "55-1_Cob") ~ str_sub(sequence, start = 30, end = 890),
                              str_detect(label, "49-2_Cob") ~ str_sub(sequence, start = 30, end = 890),
                              str_detect(label, "49-3_Cob") ~ str_sub(sequence, start = 30, end = 890),
                              str_detect(label, "49-1_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "49-2_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "49-3_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "55-1_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "55-4_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "55-5_Cox1") ~ str_sub(sequence, start = 30, end = 900),
                              str_detect(label, "49-1_cp23S") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "55-1_cp23S") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "49-2_cp23S") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "49-3_cp23S") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "49-1_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "49-2_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "49-3_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "55-1_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "55-4_LSU") ~ str_sub(sequence, start = 30, end = 500),
                              str_detect(label, "55-5_LSU") ~ str_sub(sequence, start = 30, end = 500),
           TRUE ~ str_sub(sequence, start = 30, end = 800)))



f_dss <- f_all_fasta_sub %>%
  deframe() %>%
  as_dna() %>%
  dna_to_DNAStringset()

r_dss <- r_all_fasta_sub  %>%
  deframe() %>%
  as_dna() %>%
  dna_to_DNAStringset()  %>%
  reverseComplement()
  

pair_list <- list()
for(i in 1:length(f_dss)){
  ss <- DNAStringSet(c(f_dss[i], r_dss[i]))
  pair_list[[i]] <- ss
}


# Align the sample pairs
alignment_list <- list()
for(i in 1:length(pair_list)){
  alignment <- AlignSeqs(pair_list[[i]], verbose = FALSE)
  alignment_list[[i]] <- alignment
  
   cat("Alignment", i, ":\n")
  print(alignment)
}

# View the alignments
aligned_df <- data.frame()
for(i in 1:length(alignment_list)){
  a_pair <- alignment_list[[i]] %>% writeXStringSet("temp_file.fasta")
  a_df <- read_fasta_df("temp_file.fasta")
  aligned_df <- rbind(aligned_df, a_df)
}

aligned_plotting <- aligned_df %>%
  mutate(sample_id = case_when(
    str_detect(label, "cp23S") ~ str_sub(label, 17, 26),
    str_detect(label, "Cox1|psbA") ~ str_sub(label, 17, 25),
    TRUE ~ str_sub(label, 17, 24)
  )) %>%
   mutate(label = case_when(
    str_detect(label, "cp23S") ~ str_sub(label, 34, 34),
    str_detect(label, "Cox1|psbA") ~ str_sub(label, 32, 32),
    TRUE ~ str_sub(label, 30, 30)
  )) 
  
# Create profile key
key <- aligned_plotting %>%
  tibble::rownames_to_column(var = "id")

# Create long dataframe for ggplot
long_sequences <- str_split(aligned_plotting$sequence, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(x = row_number(),
         L1 = as.character(L1)) %>%
  left_join(., key, by = c("L1" = "id")) %>%
  ungroup()

# Plot alignment
p1 <- ggplot(long_sequences, aes(y = label, x = x)) +
      geom_tile(aes(fill = value), size = 1, name = "base") +
      facet_wrap(~ sample_id, nrow = 3, scales = "free") +
      geom_vline(xintercept = seq(50, max(long_sequences$x), by = 50), color = "black", linetype = "dashed", size = 0.05) +
      scale_fill_manual(values = palette) +
      theme(aspect.ratio = 0.3,
            axis.title.y = element_blank(),
        strip.text = element_text(margin = margin(0, 0, 10, 0)), # Add space below strip text
        strip.background = element_blank(), # Remove strip background
        strip.placement = "outside",
        axis.text.x = element_text(size = 0.5),
         axis.ticks.width = unit(0.1, "cm"))+
      scale_x_continuous(expand = c(0, 0), breaks = seq(min(long_sequences$x), max(long_sequences$x), by = 100)) +
    xlab("Position")
p1

ggsave("F-R_allignment.pdf", plot = p1, device="pdf", width = 10, height = 4)

```


##Exclude list

```{r}

align_list_final <- list()
for(i in 1:length(alignment_list)){
  if(str_detect(names(alignment_list[[i]][1]), "psbA")) align_list_final[[i]] <- NULL 
  else align_list_final[[i]] <- alignment_list[[i]]
}

align_list_final <- Filter(Negate(is.null), align_list_final)


```

## Create the consensus sequences

```{r}
consensus_list <- list()
for(i in 1:length(align_list_final)){
  a_con <- align_list_final[[i]] %>% ConsensusSequence()
  names(a_con) <- names(align_list_final[[i]])[1]
  consensus_list[[i]] <- a_con
}

sequence_names <- c("SCF049-1-Cob", "SCF049-2-Cob", "SCF049-3-Cob","SCF055-1-Cob", "SCF049-1-Cox1", "SCF049-2-Cox1", "SCF049-3-Cox1","SCF055-1-Cox1", "SCF055-4-Cox1", "SCF055-5-Cox1", "SCF049-1-cp23s", "SCF049-2-cp23s", "SCF049-3-cp23s","SCF055-1-cp23s","SCF049-1-LSU", "SCF049-2-LSU", "SCF049-3-LSU","SCF055-1-LSU", "SCF055-4-LSU", "SCF055-5-LSU")
write_fasta(consensus_list, sequence_names, "consensus_sequences.fasta")
```

# View the consensus sequences

```{r}
# View the consensus
consensus_df <- data.frame()
for(i in 1:length(consensus_list)){
  a_pair <- consensus_list[[i]] %>% writeXStringSet("temp_file.fasta")
  a_df <- read_fasta_df("temp_file.fasta")
  consensus_df <- rbind(consensus_df, a_df)
}

aligned_plotting <- consensus_df %>%
  mutate(sample_id = case_when(
    str_detect(label, "cp23S") ~ str_sub(label, 17, 26),
    str_detect(label, "Cox1|psbA") ~ str_sub(label, 17, 25),
    TRUE ~ str_sub(label, 17, 24)
  )) %>%
   mutate(label = case_when(
    str_detect(label, "cp23S") ~ str_sub(label, 34, 34),
    str_detect(label, "Cox1|psbA") ~ str_sub(label, 32, 32),
    TRUE ~ str_sub(label, 30, 30)
  )) 
  
# Create profile key
key <- aligned_plotting %>%
  tibble::rownames_to_column(var = "id")

# Create long dataframe for ggplot
long_sequences <- str_split(aligned_plotting$sequence, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(x = row_number(),
         L1 = as.character(L1)) %>%
  left_join(., key, by = c("L1" = "id")) %>%
  ungroup()

p2 <- ggplot(long_sequences, aes(y = label, x = x)) +
      geom_tile(aes(fill = value), size = 1, name = "base") +
      facet_wrap(~ sample_id, nrow = 3, scales = "free") +
      geom_vline(xintercept = seq(50, max(long_sequences$x), by = 50), color = "black", linetype = "dashed", size = 0.05) +
      scale_fill_manual(values = palette) +
      theme(aspect.ratio = 0.3,
            axis.title.y = element_blank(),
        strip.text = element_text(margin = margin(0, 0, 10, 0)), # Add space below strip text
        strip.background = element_blank(), # Remove strip background
        strip.placement = "outside",
        axis.text.x = element_text(size = 0.5),
         axis.ticks.width = unit(0.1, "cm"))+
      scale_x_continuous(expand = c(0, 0), breaks = seq(min(long_sequences$x), max(long_sequences$x), by = 100)) +
    xlab("Position")
p2

ggsave("consensus_sequences.pdf", plot = p2, device="pdf", width = 10, height = 4)
```


#Treat psbA separately

```{r}
psba_F <- sort(list.files(path = "/Users/bross/Desktop/AIMS/Analysis/sanger_analysis/sanger_raw", pattern = "psbA_F.txt", full.names = TRUE))

psba_R <- sort(list.files(path = "/Users/bross/Desktop/AIMS/Analysis/sanger_analysis/sanger_raw", pattern = "psbA_R.txt", full.names = TRUE))

psba_F
psba_R

f_psba_fasta <- psba_F %>%
  map(read_fasta_df) %>% # read in all the files
  purrr::reduce(rbind) # reduce with rbind into one dataframe

r_psba_fasta <- psba_R %>%
  map(read_fasta_df) %>% # read in all the files
  purrr::reduce(rbind) # reduce with rbind into one dataframe

psbaF_fasta_sub <- f_psba_fasta %>%
  mutate(sequence = case_when(str_detect(label, "49-1") ~ str_sub(sequence, start = 40, end = 560),
                              str_detect(label, "49-2") ~ str_sub(sequence, start = 40, end = 610),
                              str_detect(label, "49-3") ~ str_sub(sequence, start = 40, end = 530),
                              TRUE ~ str_sub(sequence, start = 40, end = 600)))


psbaR_fasta_sub <- r_psba_fasta %>%
  mutate(sequence = case_when(str_detect(label, "49-1") ~ str_sub(sequence, start = 20, end = 1080),
                              str_detect(label, "49-2") ~ str_sub(sequence, start = 20, end = 520),
                              str_detect(label, "49-3") ~ str_sub(sequence, start = 20, end = 640),
           TRUE ~ str_sub(sequence, start = 20, end = 1100)))

f_psba_dss <- psbaF_fasta_sub %>%
  deframe() %>%
  as_dna() %>%
  dna_to_DNAStringset()

r_psba_dss <- psbaR_fasta_sub  %>%
  deframe() %>%
  as_dna() %>%
  dna_to_DNAStringset() %>%
  reverseComplement()

align_psba_F <- AlignSeqs(f_psba_dss)
align_psba_R <- AlignSeqs(r_psba_dss)

aligned_list <- list(align_psba_F, align_psba_R)

# View the alignments
aligned_df <- data.frame()
for(i in 1:length(aligned_list)){
  a_pair <- aligned_list[[i]] %>% writeXStringSet("temp_file.fasta")
  a_df <- read_fasta_df("temp_file.fasta")
  aligned_df <- rbind(aligned_df, a_df)
}

aligned_plotting <- aligned_df %>%
  mutate(sample_id = str_sub(label, 17, 25)) %>%
   mutate(label = str_sub(label, 32, 32))
    
# Create profile key
key <- aligned_plotting %>%
  tibble::rownames_to_column(var = "id")

key$id <- ifelse(key$id == 2, 3, ifelse(key$id == 3, 2, key$id))
key$id <- ifelse(key$id == 5, 6, ifelse(key$id == 6, 5, key$id))

# Create long dataframe for ggplot
long_sequences <- str_split(aligned_plotting$sequence, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(x = row_number(),
         L1 = as.character(L1)) %>%
  left_join(., key, by = c("L1" = "id")) %>%
  ungroup()

# Plot alignment
p3 <- ggplot(long_sequences, aes(y = sample_id, x = x)) +
      geom_tile(aes(fill = value), size = 1, name = "base") +
      facet_wrap(~ label, nrow = 3, scales = "free") +
      geom_vline(xintercept = seq(50, max(long_sequences$x), by = 50), color = "black", linetype = "dashed", size = 0.05) +
      scale_fill_manual(values = palette) +
      theme(aspect.ratio = 0.3,
            axis.title.y = element_blank(),
        strip.text = element_text(margin = margin(0, 0, 10, 0)), # Add space below strip text
        strip.background = element_blank(), # Remove strip background
        strip.placement = "outside",
        axis.text.x = element_text(size = 0.5),
         axis.ticks.width = unit(0.1, "cm"))+
      scale_x_continuous(expand = c(0, 0), breaks = seq(min(long_sequences$x), max(long_sequences$x), by = 100)) +
    xlab("Position")
p3

ggsave("psbA_allignment_049_055.pdf", plot = p3, device="pdf", width = 10, height = 4)

```


#psbA consensus 
```{r}
align_list_final <- align_psba_R[-4] #removal SCF055 sequence
consensus <- ConsensusSequence(align_list_final)
names(consensus) <- "SCF049-1-psbA"
trimmed_consensus <- subseq(consensus, start = 1050, end = 1300) #trim to the matching region only
writeXStringSet(trimmed_consensus, "consensus_psbA_049.fasta")

align_list_final2 <- align_psba_R[4] #keeping only SCF055 sequence
names(align_list_final2) <- "SCF055-1-psbA"
trimmed_consensus2 <- subseq(align_list_final2, start = 1050, end = 1300) #trim to the matching region only
writeXStringSet(trimmed_consensus2, "consensus_psbA_055.fasta")


```


#Create a unique fasta file with all the consensus sequence for all the markers
```{r}
consensus_sequences <- readDNAStringSet("consensus_sequences.fasta")
consensus_psbA_049 <- readDNAStringSet("consensus_psbA_049.fasta")
consensus_psbA_055 <- readDNAStringSet("consensus_psbA_055.fasta")

# Combine the DNAStringSets
combined_sequences <- c(consensus_sequences, consensus_psbA_049, consensus_psbA_055)
writeXStringSet(combined_sequences, filepath = "final_sequences.fasta")

```




###############################################################
#Primer design for NGS psbA variable region


#psbA consensus for primer design 
```{r}
consensus_primer_R <- ConsensusSequence(align_psba_R)
consensus_primer_F <- ConsensusSequence(align_psba_F)

names(consensus_primer_R) <- "psbA_R_Cons"
names(consensus_primer_F) <- "psbA_F_Cons"

cons_align <- c(consensus_primer_R, consensus_primer_F)
F_R_align <- AlignSeqs(cons_align)

writeXStringSet(consensus_primer_R, "variable_psbA_R.fasta")
writeXStringSet(consensus_primer_F, "variable_psbA_F.fasta")
writeXStringSet(F_R_align, "variable_psbA_F_R_align.fasta")


psbA_primer_df_R <- read_fasta_df("variable_psbA_R.fasta")
psbA_primer_df_F <- read_fasta_df("variable_psbA_F.fasta")
psbA_primer_df<- rbind(psbA_primer_df_R, psbA_primer_df_F)

psbA_primer_aligned_df <- read_fasta_df("variable_psbA_F_R_align.fasta")

#plotting for both aligned and not aligned Fwd and Rev consensus sequences. The consensus sequences have been generated by aligning Fwd with Fwd and Rev with Rev of 055 and 3 repeats of 049. No alignment between Fwd and Rev has been done. Here plotting for not aligned consensus Fwd and Rev (p4) and a further alignment between Fwd consensus and Rev consensus is shown in p5. Fwd and Rev sequences do not align, the PCR psbA amplicon should be max 1200 bp, here with the aligned sequences goes beyond this limit, which suggests issues in the alignment between Fwd and Rev. 

key <- psbA_primer_df %>%
  tibble::rownames_to_column(var = "id")

key_aligned <- psbA_primer_aligned_df %>%
  tibble::rownames_to_column(var = "id")


long_sequences <- str_split(psbA_primer_df$sequence, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(x = row_number(),
         L1 = as.character(L1)) %>%
  left_join(., key, by = c("L1" = "id")) %>%
  ungroup()

long_sequences_aligned <- str_split(psbA_primer_aligned_df$sequence, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(x = row_number(),
         L1 = as.character(L1)) %>%
  left_join(., key_aligned, by = c("L1" = "id")) %>%
  ungroup()




p4 <- ggplot(long_sequences, aes(y = label, x = x)) +
      geom_tile(aes(fill = value), size = 1, name = "base") +
      geom_vline(xintercept = seq(0,max(long_sequences$x), by = 20), color = "black", linetype = "dashed", size = 0.05) +
      scale_fill_manual(values = palette) +
      theme(aspect.ratio = 0.3,
            axis.title.y = element_blank(),
        strip.text = element_text(margin = margin(0, 0, 10, 0)), # Add space below strip text
        strip.background = element_blank(), # Remove strip background
        strip.placement = "outside",
        axis.text.x = element_text(size = 0.5),
         axis.ticks.width = unit(0.1, "cm"))+
      scale_x_continuous(expand = c(0, 0), breaks = seq(min(long_sequences$x), max(long_sequences$x), by = 10)) +
    xlab("Position")
p4

ggsave("consensuspsbA_F_R_primer.pdf", plot = p4, device="pdf", width = 10, height = 4)

p6 <- ggplot(long_sequences_aligned, aes(y = label, x = x)) +
      geom_tile(aes(fill = value), size = 1, name = "base") +
      geom_vline(xintercept = seq(0 ,max(long_sequences$x), by = 20), color = "black", linetype = "dashed", size = 0.05) +
      scale_fill_manual(values = palette) +
      theme(aspect.ratio = 0.3,
            axis.title.y = element_blank(),
        strip.text = element_text(margin = margin(0, 0, 10, 0)), # Add space below strip text
        strip.background = element_blank(), # Remove strip background
        strip.placement = "outside",
        axis.text.x = element_text(size = 0.5),
         axis.ticks.width = unit(0.1, "cm"))+
      scale_x_continuous(expand = c(0, 0), breaks = seq(min(long_sequences$x), max(long_sequences$x), by = 10)) +
    xlab("Position")
p6

ggsave("align_final_psbA.pdf", plot = p6, device="pdf", width = 10, height = 4)

```
#PCR template generation 
```{r}
trimmed_F <- subseq(consensus_primer_F, start = 1, end = 445)
trimmed_R <- subseq(consensus_primer_R, start = 928, end = 1231)

writeXStringSet(trimmed_F, filepath = "psbA_variable_PCR_template_1.fasta")
writeXStringSet(trimmed_R, filepath = "psbA_variable_PCR_template_2.fasta")

```

#Check Template 
```{r}
to_align_1 <- c(trimmed_F, consensus_primer_F)
ctrl_1 <- AlignSeqs(to_align_1) 

to_align_2 <- c(trimmed_R, consensus_primer_R)
ctrl_2 <- AlignSeqs(to_align_2) 


writeXStringSet(ctrl_1, filepath="PCR_template_align_1.fasta")
writeXStringSet(ctrl_2, filepath="PCR_template_align_2.fasta")

ctrl_df <- read_fasta_df("PCR_template_align_1.fasta")
ctrl_df2 <- read_fasta_df("PCR_template_align_2.fasta")
ctrl_df<- rbind(ctrl_df, ctrl_df2)

key <- ctrl_df %>%
  tibble::rownames_to_column(var = "id")

long_sequences <- str_split(ctrl_df$sequence, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(x = row_number(),
         L1 = as.character(L1)) %>%
  #left_join(., key, by = c("L1" = "id")) %>%
  ungroup()



p5 <- ggplot(long_sequences, aes(y = L1, x = x)) +
      geom_tile(aes(fill = value), size = 1, name = "base") +
      geom_vline(xintercept = seq(0,max(long_sequences$x), by = 50), color = "black", linetype = "dashed", size = 0.05) +
      scale_fill_manual(values = palette) +
      theme(aspect.ratio = 0.3,
            axis.title.y = element_blank(),
        strip.text = element_text(margin = margin(0, 0, 10, 0)), # Add space below strip text
        strip.background = element_blank(), # Remove strip background
        strip.placement = "outside",
        axis.text.x = element_text(size = 0.5),
         axis.ticks.width = unit(0.1, "cm"))+
      scale_x_continuous(expand = c(0, 0), breaks = seq(min(long_sequences$x), max(long_sequences$x), by = 10)) +
    xlab("Position")
p5

ggsave("PCR_template_CTRL.pdf", plot = p5, device="pdf", width = 10, height = 4)


```












#Treat psbA ab1 files - not used since I have txt but for further need. This code read txt and ab1 files, pull the sequences, merge them together, align and plot.

```{r}
ab1_for <- list.files(path = "/Users/bross/Desktop/AIMS/Analysis/sanger_analysis/psba_055", pattern = "\\F.ab1$", full.names = TRUE)

ab1_rev <- list.files(path = "/Users/bross/Desktop/AIMS/Analysis/sanger_analysis/psba_055", pattern = "\\R.ab1$", full.names = TRUE)

read_ab1_df <- function(file) {
    ab1_data <- read.abif(file)
    if (inherits(ab1_data, "abif")) {
        seq <- as.character(sangerseqR::primarySeq(sangerseqR::sangerseq(ab1_data)))
        return(data.frame(label = basename(file), sequence = seq, stringsAsFactors = FALSE))
    } else {
        stop("The file does not seem to be a valid AB1 file.")
    }
}

f_ab1_fasta <- ab1_for %>%
  map(read_ab1_df) %>%
  reduce(rbind)

r_ab1_fasta <- ab1_rev %>%
  map(read_ab1_df) %>%
  reduce(rbind)

combined_fasta <- bind_rows(f_ab1_fasta, r_ab1_fasta)


psbaF_fasta_sub <- combined_fasta_f %>%
  mutate(sequence = case_when(str_detect(label, "49-1") ~ str_sub(sequence, start = 40, end = 560),
                              str_detect(label, "49-2") ~ str_sub(sequence, start = 40, end = 610),
                              str_detect(label, "49-3") ~ str_sub(sequence, start = 40, end = 530),
                              TRUE ~ str_sub(sequence, start = 30, end = 600)))


psbaR_fasta_sub <- combined_fasta_r%>%
  mutate(sequence = case_when(str_detect(label, "49-1") ~ str_sub(sequence, start = 20, end = 1080),
                              str_detect(label, "49-2") ~ str_sub(sequence, start = 20, end = 520),
                              str_detect(label, "49-3") ~ str_sub(sequence, start = 20, end = 640),
           TRUE ~ str_sub(sequence, start = 30, end = 800)))

f_psba_ab1 <- combined_fasta %>%
  deframe() %>%
  as_dna() %>%
  dna_to_DNAStringset() %>%
  reverseComplement() %>%
  AlignSeqs()

aligned_df_55 <- data.frame(
  label = names(f_psba_ab1),
  sequence = as.character(f_psba_ab1),
  stringsAsFactors = FALSE
)

aligned_plotting_2 <- aligned_df_55 %>%
  mutate(sample_id = str_sub(label, 17, 25)) %>%
   mutate(label = str_sub(label, 32, 32)) %>%
  rownames_to_column(var = "row_number")

    
# Create profile key
key <- aligned_plotting_2 %>%
  tibble::rownames_to_column(var = "id")

# Create long dataframe for ggplot
long_sequences <- str_split(aligned_plotting_2$sequence, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(x = row_number(),
         L1 = as.character(L1)) %>%
  left_join(., key, by = c("L1" = "id")) %>%
  ungroup()

# Plot alignment
p3 <- ggplot(long_sequences, aes(y = label, x = x)) +
  geom_tile(aes(fill = value), size = 1) +
  geom_vline(xintercept = seq(50, max(long_sequences$x), by = 50), color = "black", linetype = "dashed", size = 0.05) +
  scale_fill_manual(values = palette, name = "base") +
  theme(
    aspect.ratio = 0.3,
    axis.title.y = element_blank(),
    strip.text = element_text(margin = margin(0, 0, 10, 0)), # Add space below strip text
    strip.background = element_blank(), # Remove strip background
    strip.placement = "outside",
    axis.text.x = element_text(size = 0.5),
    axis.ticks.length = unit(0.1, "cm")
  ) +
  scale_x_continuous(expand = c(0, 0), breaks = seq(min(long_sequences$x), max(long_sequences$x), by = 100)) +
  xlab("Position")

p3


ggsave("psbA_allignment_055.pdf", plot = p3, device="pdf", width = 10, height = 4)


aligned_plotting_2 <- aligned_plotting_2 %>% select(-row_number)
joined <- rbind(aligned_plotting, aligned_plotting_2)


key <- joined %>%
  tibble::rownames_to_column(var = "id")

key$id <- ifelse(key$id == 2, 3, ifelse(key$id == 3, 2, key$id))
key$id <- ifelse(key$id == 5, 6, ifelse(key$id == 6, 5, key$id))

# Create long dataframe for ggplot
long_sequences <- str_split(joined$sequence, "") %>%
  reshape2::melt() %>%
  group_by(L1) %>%
  mutate(x = row_number(),
         L1 = as.character(L1)) %>%
  left_join(., key, by = c("L1" = "id")) %>%
  ungroup()

# Plot alignment
p3 <- ggplot(long_sequences, aes(y = sample_id, x = x)) +
      geom_tile(aes(fill = value), size = 1, name = "base") +
      facet_wrap(~ label, nrow = 3, scales = "free") +
      geom_vline(xintercept = seq(50, max(long_sequences$x), by = 50), color = "black", linetype = "dashed", size = 0.05) +
      scale_fill_manual(values = palette) +
      theme(aspect.ratio = 0.3,
            axis.title.y = element_blank(),
        strip.text = element_text(margin = margin(0, 0, 10, 0)), # Add space below strip text
        strip.background = element_blank(), # Remove strip background
        strip.placement = "outside",
        axis.text.x = element_text(size = 0.5),
         axis.ticks.width = unit(0.1, "cm"))+
      scale_x_continuous(expand = c(0, 0), breaks = seq(min(long_sequences$x), max(long_sequences$x), by = 100)) +
    xlab("Position")
p3

```


